\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2017-01-11 doc/publications/plos16/plos16.tex
Rewrite RDD section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest.}
VariantSpark made use of Spark's machine learning algorithms ``Spark ML''. These algorithms require data in a
``DataFrame'', similar to DataFrames in R.  While Spark ML algorithms demonstrate some scalability when dealing with a
large number of samples, this scalability eventually break down as we include more features.

One of the restrictions preventing Spark ML algorithms from scaling to high-dimensionality data is that the feature-set
for each sample is stored as a vector data type. Spark vectors are stored in resilient distributed dataset (RDDs), and
while RDDs can be split, vectors cannot. So as we add more samples (in our case individuals), Spark ML can simply split
the RDD into more partitions and resource requirements will remain relatively stable. However, as the number of
dimensions (variants) increases, the size of the vectors increases. This results in higher memory-requirements as
increasingly large vectors must be loaded into memory.  \marginpar{not clear. Are RDDs the problem or the solutuion}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest}
VariantSpark made use of Spark's machine learning algorithms ``Spark ML''. While Spark ML algorithms demonstrate
scalability when dealing with a large number of samples, this scalability eventually breaks down as we include more
features.

As is standard with Spark applications, we store our data in a Resilient Distributed Dataset (RDD), where an RDD is
essentially a collection of elements. In the case of Spark ML, each element in our RDD is a sample. RDDs contribute to
the scalability of Spark as they can distributed across multiple nodes and operated on in parallel. Even as we add more
samples to a dataset, Spark can still simply schedule extra tasks to handle the additional items in the RDD.

However, within an RDD, spark ML stores each sample as a vector. Unlike RDDs, which can be partitioned and distributed
across multiple nodes, each vector must be present in its entirity on any node accessing it. This is no problem with
typical datasets, however as dimensionality increases, the vectors eventually reach a size where they can no longer fit
into a single node's memory.

So in the case of adding more samples, Spark ML can simply create more tasks, keeping memory consumption within the
cluster's bounds. However, as the dimensionality of each sample grows, the memory requirements of the job increases to
enable these increasingly large vectors to be loaded into memory.
\marginpar{not clear. Are RDDs the problem or the solutuion Aidan: Better?}
